### 2021-02-23 test.Rmd
Add files via upload
Draft of article for submission to the Journal of Integrated Pest Management, Category: 'Issues'

---
title: "test"
author: "Austin N. Fife"
date: "2/22/2021"
output: 
  word_document:
    reference_docx: ../r-templates/word-template-jis-01.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.