### 2021-06-10 ofv_fife_2021.Rmd
New citations

---
title: "For submission to the Journal of Integrated Pest Management"
subtitle: "Category: 'Brief Communications'"
output: 
  word_document:
    reference_docx: ../r-word-templates/word-template-jis.docx
# - html_document
bibliography: 'bib/thesis.bib'
citation_package: natbib
biblio-style: apalike
csl: 'entsoc.csl'
#always_allow_html: true
---
```{r setup, include=FALSE}
#setup options
knitr::opts_chunk$set(fig.width=12, fig.height=8, echo=FALSE)

#a list of packages used for scripts
pkgs <- c('captioner', 'pryr', 'tiff')

#loading required packages
lapply(pkgs, library, character.only = TRUE)
rm(pkgs)

#setting up captions for use in the document
fig_nums <- captioner::captioner(prefix = 'Fig.')
tab_nums <- captioner::captioner(prefix = 'Table')

#allows me to use a custom function instead of the citation function of captioner, which is wordy to use
citef <- pryr::partial(fig_nums, display = "cite")
citet <- pryr::partial(tab_nums, display = "cite")
```

### First report of the *Brevipalpus*-transmitted (Trombidiformes: Tenuipalpidae) *Orchid fleck dichorhavirus* infecting three ornamentals in Florida

Austin **Fife**^1^, Daniel **Carrillo**^2^, Gary **Knox**^3^, Fanny **Iriarte**^4^, Kishore **Dey**^5^, Avijit **Roy**^6^, Ronald **Ochoa**^7^, Gary **Bauchan**^8^$\dagger$, Mathews **Paret**^4,9^, Xavier **Martini**^1^\*

^1^ University of Florida, Department of Entomology and Nematology, North Florida Research and Education Center, Quincy FL 32351

^2^ University of Florida, Department of Entomology and Nematology, Tropical Research and Education Center, Homestead FL 33031

^3^ University of Florida, Department of Environmental Horticulture, North Florida Research and Education Center, Quincy FL 32351

^4^ University of Florida, Department of Plant Pathology, North Florida Research and Education Center, Quincy FL 32351

^5^ The Florida Department of Agriculture and Consumer Services, Division of Plant Industry, Section of Plant Pathology, Doyle Conner Building, 1911 SW 34th street, Gainesville, FL 32608

^6^	United States Department of Agriculture – Agriculture Research Service, Molecular Plant Pathology Laboratory, 10300 Baltimore Ave, Bldg. 4 BARC-West, Beltsville, MD 20705

^7^ United States Department of Agriculture - Agriculture Research Service, Systematic Entomology Laboratory 10300 Baltimore Ave, Bldg. 5 BARC-West, Beltsville, MD 20705

^8^	United States Department of Agriculture - Animal and Plant Health Inspection Service, Electron and Confocal Microscopy Unit, Bldg. 12  BARC-West, 10300 Baltimore Ave, Beltsville, MD 20705

^9^ Plant Pathology Department, University of Florida, Gainesville, FL 32611

*Corresponding author; E-mail: xmartini\@ufl.edu Phone: 850-875-7160 Fax: 352-846-6617

\pagebreak

### Abstract

<!-- 250 words or less. -->
<!-- Give scientific name and authority at first mention of each organism. -->
<!-- Do not cite references, figures, tables, probability levels, or results. -->
<!-- Refer to results only in the general sense. -->
<!-- A second abstract in a second language is permitted. -->

We describe the first outbreaks of *Orchid fleck dichorhavirus*, belonging to the orchid-infecting subgroup (OFV-Orc), from three unreported hosts: *Liriope muscari*, cv. 'Gigantea' (Decaisne) Bailey, *Ophiopogon intermedius* Don and *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) in Leon and Alachua Counties, FL. Strains of OFV-Orc infect over 50 plant species belonging to the plant families Orchidaceae, Asparagaceae (Nolinoidaea), and infects *Citrus* (Rutaceae) as citrus leprosis disease. The only known vectors of OFV-Orc are flat mites, genus *Brevipalpus* (Trombidiformes: Tenuipalpidae). Florida has various plants in the landscape which *Brevipalpus* spp. feed on, which are susceptible to infection by OFV-Orc.

Chlorotic ringspots and flecking were seen affecting Liriopogons (*Liriope* and *Ophiopogon* spp.) in Leon County, FL. Nearby *A. elatior* also appeared chlorotic. Local diagnostics returned negative for common plant pathogens, therefore new samples were sent to the Florida Department of Agriculture and Consumer Services (FDACS) and USDA-ARS for identification.

Two orchid-infecting strains of Orchid fleck virus were detected via combinations of conventional RT-PCR, RT-qPCR, Sanger sequencing and High Throughput Sequencing. Amplicons shared 98% nucleotide identity with OFV-Orc1 and OFV-Orc2 available in NCBI GenBank. Coinfections were seen in each county, but single strains of OFV-Orc were seen in *L. muscari* (Alachua, OFV-Orc2) and *A. elatior* (Leon, OFV-Orc1).

Three potential mite vectors were identified via cryo-scanning electron microscopy (Cryo-SEM): *Brevipalpus californicus* (Banks) sensu lato, *B. obovatus* Donnadieu, and *B. confusus* Baker.

In conclusion, *Orchid fleck dichorhavirus* is present in northern Florida, representing a risk for susceptible plants in the southeastern US.

### Keywords:
False spider mite, flat mite, *Brevipalpus*-transmitted viruses, *Liriope*, Nolinoidaea, *Ophiopogon*, *Aspidistra*, Ruscaceae, Rutaceae, Asparagaceae, orchid, Orchidaceae, pests, ornamental plants, Orchid fleck virus.

\pagebreak
*Orchid fleck dichorhavirus*, commonly referred to as Orchid Fleck Virus (OFV), is the type member for the genus *Dichorhavirus*, family *Rhabdoviridae*. OFV is a bacilliform, nuclear rhabdovirus composed of two segments of single-stranded, negative-sense RNA which infects plants [@Amarasinghe2019; @Dietzgen2014; @Walker2018]. Flat mites from the genus *Brevipalpus* (Trombidiformes: Tenuipalpidae) are the only group known to transmit dichorhaviruses [@Maeda1998], and *Brevipalpus californicus* (Banks) sensu lato are the only mites which do so in a persistent propagative manner [@Kondo2003].

OFV-infected plants exhibit various symptoms depending on the infected plant species as well as the strain of the OFV associated with the infection [@Kubo2009b], but symptoms typically appear as chlorotic flecks, which ultimately coalesce into ringspot patterns.

OFV was first described as infecting *Cymbidium* orchids in Japan [@Doi1977]. OFV and OFV-like rhabdoviruses have been reported infecting orchids in Asia, Africa, North America, South America, Europe, and Oceania. The prevalence of OFV and its mite vector is thought to be associated with the movement of infected orchids [@Dietzgen2018a].

OFV naturally infects more than fifty species of Orchidaceae [@Kitajima2010; @Peng2013], some Asparagaceae (Nolinoidaea) [@Mei2016; @Dietzgen2018b], and Rutaceae, where it causes citrus leprosis-like symptoms [@Cook2019; @Velarde2021; @Roy2015; @Roy2020]. Mechanical transmission of OFV is possible under laboratory conditions to some plants belonging to the plant families Chenopodiaceae, Aizoaceae, Fabaceae, and Solanaceae [@Chang1976; @Kondo2003; @Peng2013].

During June 2020, chlorotic ringspot symptoms were observed on Giant Lilyturf *Liriope* spp., cv. 'Gigantea' in a landscape of Leon County, Florida (`r citef("fig_1")`). *Liriope* belong to a group of plants in the family Asparagaceae, subfamily Nolinoidaea, comprised of grass-like monocotyledonous liliod plants native to southeastern Asia [@Chase2009; @Meng2021]. *Liriope* and the closely related *Ophiopogon* (Asparagaceae: Nolinoidaea) are considered the most important ground cover plant in the southeastern United States [@Mcharo2003].

Viral infections of suspected leaf samples were initially tested at the Plant Disease Diagnostic Clinic at the North Florida Research and Education Center (NFREC) in Quincy, FL. All the samples were tested with one step conventional RT-PCR, and were found negative for begomovirus, carlavirus, potyvirus, tospovirus, Cucumber mosaic virus and Tobacco mosaic virus.

As initial diagnostics were inconclusive, new samples were collected during July and August of 2020 to collect more of these putatively infected plants with ringspot symptoms. The plants collected included *Liriope* spp. and *Ophiopogon* spp., as well as *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea). *A. elatior* was suspected to be infected, due to both its proximity to infected *Liriope* and the presence of unusually chlorotic leaves (`r citef("fig_2")`). Upon collection, the new samples were sent to the Florida Department of Agriculture and Consumer Services (FDACS) for identification.

The FDACS determined that the pathogen was *Orchid fleck dichorhavirus* using previously published primers and methods to conduct RT-PCR and Sanger sequencing [@Kubo2006a; @Kubo2006b; @Kubo2009a; @Kubo2009b; @RamosGonzalez2015]. Orchid subgroup 1, OFV-Orc was identified following the methods described in @Kondo2017. Sequencing demonstrated a shared 98\% nucleotide identity with the orchid strain subgroup, OFV-Orc (isolates So and Br with GenBank Accession numbers: AB244418 and MK522807, respectively) [@Kondo2006; @Kondo2017].

These samples from FDACS were subsequently retested by the USDA-ARS, in conjunction with tests of fresh samples from both Alachua and Leon counties. The USDA used RT-PCR, RT-qPCR, and High Throughput Sequencing (HTS) in sequence to reconfirm the presence of *Orchid fleck dichorhavirus*. RT-PCR and qPCR with Generic R2-Dicho-GF and R2-Dicho-GR primers amplifed ~800 nt of L-gene (RNA2) amplicon [@Roy2020], and OFV-Orc1 and OFV-Orc2 were detected in both *O. intermedius* and *A. elatior* from Leon County.

HTS reaffirmed the presence of OFV-Orc1 and OFV-Orc2 strains in Leon and Alachua counties (`r citet("tab_1")`). HTS results from Leon County revealed that *L. muscari* were coinfected with both strains (OFV-Orc1 and OFV-Orc2), while *A. elatior* were solely infected with OFV-Orc1. HTS of *L. muscari* from Alachua County revealed infections with the OFV-Orc2 strain.

After the initial identification by FDACS of OFV-Orc strains, mite samples were collected from symptomatic Asparagaceae in Leon County. Most mites collected were Tenuipalpid mites (flat mites or false spider mites), a known pest of ornamental plants, some of which are known to act as vectors for plant viruses [@Childers2003; @Childers2011].

Mite taxonomy is complicated by cryptic species complexes which occur in many plant-feeding groups of the Acari [@Arthur2011; @Skoracka2010; @Skoracka2013; @Umina1999], including tenuipalpid mites from the genus *Brevipalpus* [@Navia2013]. The commonly used phase-contrast microscopy is insufficient to detect some diagnostic characters for seperation of cryptic species, instead best practices recommend the combination of Differential Interference Contrast (DIC) Microscopy and Scanning Electron Microscopy along with molecular methods to separate cryptic species [@Beard2015].

The flat mites collected were initially suspected to belong to *B. californicus* after inspection with phase contrast microscopy. Subsequent observation via DIC microscopy at FDACS agreed with this tentative identification. Unfortunately, the *B. californicus* s.l. species group, sensu @Baker1987 is suspected to contain cryptic species [@Childers2011; @Rodrigues2013]. New mite samples were collected from symptomatic liriopogons and *A. elatior* in Leon County and sent to USDA-ARS's Electron and Confocal Microscopy Unit for analysis. Three mite species were recovered and examined under cryo-scanning electron microscopy (Cryo-SEM): *B. californicus* s.l. (`r citef("fig_3")`), *B. obovatus* Donnadieu and *B. confusus* Baker (`r citef("fig_4")`).

The first report of OFV in the US is thought to be @Ko1985, who describes nuclear inclusions caused by an undescribed bacilliform rhabdovirus in *Brassia* orchids. The significance of this report is their description of the spoke-wheel configurations of the viral particles [@Ko1985], a sign typically associated with OFV infection [@Chang1976]. Unfortunately, this article made no mention of mites or further investigations of the virus. The first report of OFV in the continental US was @Bratsch2015, who confirmed the presence of OFV in *Phalaenopsis* hybrids using Transmission Electron Microscopy of ultrathin sections of plant tissue as well as molecular sequence analysis. They also discuss the association of OFV with *Brevipalpus* mites, but the authors did not make a conclusive species identification beyond suggesting that the mite vector belonged to the *B. californicus* group, referring to @Kondo2003's publication [@Bratsch2015].

Later reports of OFV described OFV infecting a previously undescribed Nolinoidaea hosts in Australia [@Dietzgen2018b; @Mei2016], including *Liriope spicata* (Thunb.) Lour, a different species of liriopogon than those identified from the Florida sites. We are not aware of any reports of OFV infecting liriopogons, *A. elatior* nor other Nolinoidaea in the US. Although @Zheng2013 had mentioned an association between *B. californicus* and *A. elatior*, they never reported symptoms of OFV-Orc in this plant. Taking this information into consideration, we believe that our findings indicate the first report of OFV-Orc infecting ornamental Nolinoidaea in Florida, and possibly the US. This publication also marks the first reports of *A. elatior* and *Ophiopogon* spp. as natural hosts of OFV-Orc.

OFV consists of two orchid strains (OFV-Orc1 and OFV-Orc2) and two citrus strains (OFV-Cit1 and OFV-Cit2) [@Roy2020; @Beltran-Beltran2020]. The OFV strains detected in Florida are identical in gene order, content, and genome sequence to the orchid strains of OFV infecting citrus in Hawaii, Mexico, Colombia, and South Africa [@Roy2020; @Beltran-Beltran2020]. Both OFV-Orc1 and OFV-Orc2 infect citrus [@Roy2020], but none of the citrus strains have been reported from any orchid species. The *Brevipalpus* mites collected from liriopogons and *A. elatior* in Leon County were abundant on OFV-infected plants very near to citrus trees, some plants even surrounding the trunk. *B. californicus* s. l. has been reported as a pest of citrus [@Childers2003] and are often collected from citrus rinds [@Baker1949; @Baker1987], so the proximity of these mite vectors to citrus raises the question: why these trees are not currently infected with OFV-Orc? Each possibility suggests further research, but there are a few possible explanations. It is important to note the uncertainty surrounding the vector for OFV-Orc. There are three mite species which have been recovered from OFV-Orc infected plants: *B. californicus* (the most likely culprit), *B. obovatus*, and *B. confusus*. Each species has its own unique biology, and all have been implicated with a variety of different hosts. This suggests that the spread of OFV-Orc would be a function of various combinations of several potential factors, possibly including host preferences, vectorial capacity, viral propagation/circulation in the vector, viral acquisition times, and feeding times required for transmission. @Ferreira2020 notes  Another possibility is that the *B. californicus* which we find on liriopogons and *A. elatior* are not actually the same species as those found on citrus, and instead represent 2-3 different cryptic species.

Best practices for integrated pest management have not been created for controlling *Brevipalpus* mites on these ornamentals, but methods designed to control other mite species may be applicable. The most common method used to control mites are synthetic acaricides and horticultural oils [@Leeuwen2015; @Gill2002]. Unfortunately, some acaricides and their residues can harm beneficial predatory mites as well [@Fernandez2017], even at low doses [@Havasi2021], and mixing different chemistries can be detrimental for mite control [@Vechia2018]. In addition, pesticide resistance has been reported in various *Brevipalpus* populations [@Alves2000; @Campos2002; @Omoto2000], possibly as the result of exposure to chemistry meant to control other arthropod pests [@Vechia2021]. Various biological control methods have been studied as well, but their use has not been widely adopted for various reasons [@Messing2017], but predatory mites [@Chen2006], entomopathogenic fungi [@Reddy2001; @Midthassel2016], host plant resistance [@Leeuwen2009], and even ultraviolet radiation [@Osakabe2021] have shown some promise for controlling other plant-feeding mites. It is often possible to integrate different biological control techniques for improved control, such as combining predatory mites with entopathogenic fungi [@Reddy2001; @Midthassel2016]. However, it is still necessary to consider the biology of biological agents to avoid incompatibilities, such as cannibalism [@Cakmak2009], requiring refuges [@Smith1991], or reapplying after conventional pesticides are used [@Smith1991]. The predatory mite *Galendromus helveolus* (Chant) (Acari: Phytoseiidae) have been reported to feed on eggs, larvae and nymphs of *B. californicus* [@Chen2006] and studies of *Amblyseius largoensis* (Muma) (Acari: Phytoseiidae) demonstrated its ability to suppress populations of *B. yothersi* on citrus plants [@Argolo2020]. In addition, the entomopathogenic fungus *Metarhizium anisopliae* var. *acridum* was found to be pathogenic to *B. phoenicis* [@Magalhaes2005].

In conclusion, detecting OFV in Florida represents a concern for horticulturists who grow orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagaceae species which are commonly used in landscaping. Florida is also home to a plethora of native and naturalized orchid species, many of which are threatened, including cultivated *Vanilla* in southern Florida [@Chambers2019] and the famous Ghost Orchid, [*Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe]. Citrus leprosis was present in Florida during the 1860’s and almost eradicated by the mid-1960s [@Childers2003; @Knorr1968a; @Knorr1968b]. An examination of herbarium specimens of Florida citrus found that this historical virus, Citrus leprosis dichorhavirus-N0, is distantly related to the modern strains of OFV [@Hartung2015; @Kitajima2011a; @Roy2020]. The recent detection of OFV-Orc1 in South Africa [@Cook2019] in *C. sinensis* (Navel and Valencia orange) and OFV-Orc2 in Hawaii [@Velarde2021] in *C. reticulata* (mandarin) and *C. jambhiri* (rough lemon) associated with leprosis-like symptoms highlights the threat of different strains of OFV on citrus, which will be a definite concern to the US multi-billion-dollar citrus industry. *B. californicus*, as well as *B. yothersi* (Baker), are both known vectors of dichorhaviruses (OFV) [@GarciaEscamilla2018; @Kondo2003; @Beltran-Beltran2020] and *B. obovatus* is a suspected vector as well [@Childers2003]. All three mite species/complexes are present in Florida [@Childers2003; @Akyazi2017] (`r citef("fig_4")`), and are difficult to identify by non-experts, or without advanced methodologies. DNA barcoding [@Armstrong2005] or a similarly simple and accurate method for identification of these mite complexes is vital to determine which of these species are responsible for transmission of OFV-Orc, and therefore which mite populations need to be monitored or controlled. By doing so, we can determine the risk OFV-Orc represents for the native plants, agriculture and the ornamental/landscaping industries of Florida and the surrounding regions.

### Acknowledgements
We would like to give special thanks to the Tallahassee Museum for their patience, cooperation, and support with collecting plant samples. We also want to thank Drs. Sam Bolton, FDACS and Aline Tassi, Univ. of Sao Paulo, Brazil for checking the mites we have sent for species validation. Furthermore, we are grateful for Dr. Marc S. Frank's identification of the Liriopogons collected. We are especially indebted to the late Dr. Gary Bauchan for his contributions to this study and the field of acarology, he will be greatly missed. This research was partly funded by the USDA National Institute of Food and Agriculture, Hatch project FLA-NFC-005607. Mention of trade names or commercial products in this publication is solely for the purpose of providing specific information and does not imply recommendation or endorsement by the USDA; USDA is an equal opportunity provider and employer.

### References
<div id="refs"></div>

### Table 1: List of Asparagaceae (Nolinoidaea) species with verified cases of *Orchid fleck dichorhavirus*, collected from the landscape of northern Florida

|Scientific Name|Common Names|County|Strains|
|:-----------|:--------------|:-----|:------|
|*Liriope muscari* cv. 'Gigantea'\* (Decaisne) Bailey | Lilyturf, Orchardgrass, Monkeygrass|Alachua & Leon|OFV-Orc1 & OFV-Orc2|
|*Ophiopogon intermedius*\*\* Don| Aztec Grass, 'Argenteomarginatus'|Leon|OFV-Orc1 & OFV-Orc2|
|*Aspidistra elatior* Blume| Cast Iron Plant, Bar-room Plant|Leon|OFV-Orc1 & OFV-Orc2|
`r tab_nums(name = "tab_1", caption = "* *Liriope muscari* cv. 'Gigantea' has been traditionally classified as *L. gigantea* Hume by @Broussard2007 and @Fantz2015, although this distinction has been challenged by @Wang2014 and @Masiero2020. * * *O. intermedius* is sometimes misclassified as *Liriope muscari* ‘Variegated Evergreen Giant’ [@Fantz2008b, @Fantz2009] or ‘Grandiflora White’ [@Fantz2009].")`

### Figure captions
`r fig_nums(name = "fig_1", caption = "Variety of symptoms expressed by *Liriope* spp. infected with *Orchid fleck dichorhavirus*: (a) ringspot symptoms on *Liriope gigantea* (b-c) Details of ringspot symptoms on *Liriope gigantea* (d) chlorotic ringspot *Liriope muscari* cv. 'Silvery Sunproof'")`

`r fig_nums(name = "fig_2", caption = "Symptoms expressed by *Aspidistra elatior* infected with *Orchid fleck dichorhavirus*: (a) Detail of leaf chlorosis (b) Chlorosis appears similar to sun damage (c-d) Chlorotic ringspot may indicate early symptoms of OFV")`

`r fig_nums(name = "fig_3", caption = "Cryo-SEM images of *Brevipalpus californicus* sensu lato displaying various characters used for identification [@Baker1987; @Beard2015] (a) Dorsum  (b) Lateral view (c) Venter (d) Close up of distal end of leg 2, with arrows indicating paired solenidia, characteristic of the genus *Brevipalpus* (e) Enlargement of the microplates of the mite cerotegument (f) Dorsal view of the distal portion of mite abdomen (g) Dorsal view of the mite rostrum (h) Ventral view of mite rostrum, observe 3 distal setae.")`

`r fig_nums(name = "fig_4", caption = "Florida is home to other common pest species of *Brevipalpus*, which are potential vectors of *Orchid fleck dichorhavirus*: (a) *B. phoenicis*, dorsal (b) *B. yothersi*, lateral (c) *B. obovatus*, dorsal.")`

### Figures
```{r, fig_1, warning=FALSE}
grid::grid.raster(tiff::readTIFF("figure/symp-ofv-lir.tif"))
```

```{r, fig_2, warning=FALSE}
grid::grid.raster(tiff::readTIFF("figure/symp-ofv-asp.tif"))
```

```{r, fig_3, warning=FALSE}
grid::grid.raster(tiff::readTIFF("figure/b_cali.tif"))
```

```{r, fig_4, warning=FALSE}
grid::grid.raster(tiff::readTIFF("figure/other_mites.tif"))
```